On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-10-06 paper.tex
Add Mathai citations
Add Mathai citations

The problems studied in this paper are related to Buffon's needle problem and Laplace's extension of it. Buffon's original 1733 problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a height $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation. For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$. This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might multiple lines.

The Buffon-Laplace needle problem considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$. The number of visited tiles equals one plus the number of crossings with probability 1, as shown in Proposition \ref{prop: 
i+j-1}. The probability of the needle staying within a single rectangle, that is visiting only one tile, is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$. This work considers the complementary question of the chance that the needle visits the maximum number of tiles possible for its length, described by a formula in Theorem~\ref{theo: probmax, sq}.

@@

\begin{proof}
First, suppose the grid is simplified to just horizontal lines a height of $b$ apart. This matches the setup of the 1733 Buffon's needle problem, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected (i.e., average) number of lines crossed equals
$$ \frac{2\len}{\pi b}.$$
Where $b<\len$, this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.

>>>>

The problems studied in this paper are related to Buffon's needle problem and Laplace's extension of it. Buffon's original 1733 problem considers tossing a needle of length $\len$ onto ruled paper with parallel lines a height $b$ apart, and asks what is the probability of the needle crossing a line if it lands in a uniformly random position and orientation. For a needle of length $\len = b$, this is $2/\pi$, and hence repeated trials of this experiment can be used to estimate $\pi$. This is generalized in \cite{Ramaley69} to $\frac{2\len}{\pi b}$ for the expected number of crossings of a needle which might intersect multiple lines.

The Buffon-Laplace needle problem, described in \cite{Mathai99}, considers a needle tossed onto a grid of rectangular tiles of width $a$ and height $b$. The number of visited tiles equals one plus the number of crossings with probability 1, as shown in Proposition \ref{prop: 
i+j-1}. The probability of the needle staying within a single rectangle, that is visiting only one tile, is computed in \cite{Arnow94} for the case where $\len < \min(a,b)$. This work considers the complementary question of the chance that the needle visits the maximum number of tiles possible for its length, described by a formula in Theorem~\ref{theo: probmax, sq}.

@@

\begin{proof}
First, suppose the grid is simplified to just horizontal lines a height of $b$ apart. This matches the setup of the 1733 Buffon's needle problem described in \cite{Mathai99}, except that here the length $\len$ of the needle may exceed the spacing $b$, allowing it to cross multiple lines. As shown in \cite{Ramaley69}, the expected (i.e., average) number of lines crossed equals
$$ \frac{2\len}{\pi b}.$$
Where $b<\len$, this expectation matches the probability of the needle crossing a grid line, since it cannot cross more than one.
